---
title: "prac1_exercises"
format:html
---
library(tidyverse)
library(gt)
Round pi to the nearest whole number
```{r}
round_pi<-round(3.14156)
print(round_pi)
```

Round to two decimal places through the digits argument
```{r}
round(3.14156,digits=2)
```

Sequence of numbers, with the arguments in the following order: from, to and by
```{r}
print(seq(1, 10, 2))
```

Rep function repeats the numbers
```{r}
print(rep(5, 7))
print(rep(x=5, times=7))
```

Variables:
```{r}
whatsmynumber <- 1
whatsmynumber <- 5
```
The variable whatsmynumber now corresponds to the number 5 

```{r}
char_var<-"MathBio"
num_var<-15
log_var<-TRUE
print(paste("My character variable is",char_var))
print(paste("My numeric variable is",num_var))
print(paste("My logical variable is",log_var))
```

Experiment to test radiation effect analysis
```{r}
CellID1<-"CX1"
CellID2<-"CX2"
TreatCell1_bool<-TRUE
TreatCell2_bool<-FALSE
NumBefore1<-5415
NumAfter1<-1432
NumBefore2<-8123
NumAfter2<-7543

```

Quirks of R:
```{r}
print(TRUE + TRUE) #Answer is 2
print(TRUE + FALSE) # Answer is 1
print(5 + TRUE) #Answer is 6

x <- NA
print(x == NA)
print(is.na(x)) #Answer is TRUE

x <- 0 / 0 
print(x) #Answer is NaN
print(is.na(x)) #Answer is TRUE
print(is.nan(x)) #Answer is TRUE

print(round(2.5)) #Round does round to even, so 2.5 becomes 2 and 3.5 becomes 4
print(round(3.5)) 

```

Part 4

Cell experiment dataset
```{r}

CellIds<-c("CX1","CX2")
Treatment<-c(TRUE,FALSE)
n_cells_before<-c(5415,8123)
n_cells_after<-c(1432,7543)
```

Isolate fourth variable
```{r}
all_my_pets <- c("dog", "horse", "guinea pig", "chicken", "cat")
fourth_pet<-all_my_pets[4]
print(fourth_pet)
```

Vectorisation
```{r}
x <- c(2, -4, 6, -2)
x * 2 #gives 4,-8,12,-4
x > 5 # gives FALSE FALSE TRUE FALSE
sum(x)
abs(x)
```

Dataframes
```{r}
library(tidyverse)
library(gt)
# You should already have this part
cell_id_1 <- "CX1"
cell_id_2 <- "CX2"

treated_1 <- TRUE
treated_2 <- FALSE

n_before_1 <- 5415
n_before_2 <- 8123

n_after_1 <- 1432
n_after_2 <- 7543

# Combine these into vectors
cell_id <- c(cell_id_1, cell_id_2)
treated <- c(treated_1, treated_2)
n_before <- c(n_before_1, n_before_2)
n_after <- c(n_after_1, n_after_2)

ExperimentData<-tibble(
  cellID=cell_id,
  treatment=treated,
  numBefore=n_before,
  numAfter=n_after
)
gt(ExperimentData)
```

We use UKHSA data on rhinovirus infections over time.
```{r}
library(tidyverse)
library(gt)
#weather <- read_csv("data/weather_data.csv")

rhinovirus <- read_csv("data/ukhsa-chart-download-rhinovirus.csv")

gt(head(rhinovirus))
```

Import data via API
```{r}
library("httr2")
url <- "https://api.ukhsa-dashboard.data.gov.uk/themes/infectious_disease/sub_themes/gastrointestinal/topics/C-difficile/geography_types/Nation/geographies/England/metrics/c-difficile_cases_countsByOnsetType"
url <- paste0(url, "?page_size=365")
print(url)
api_request <- request(url)
response <- req_perform(api_request)
json_response <- resp_body_json(response) #read json object
str(json_response, max.level = 1)
results_df <- bind_rows(json_response$results)
gt(head(results_df))
print(unique(pull(results_df, stratum)))
print(unique(pull(results_df, sex)))
print(unique(pull(results_df, age)))
filt <- filter(results_df, stratum == "Total cases")
gt(filt)

# Assign URL
url <- "https://api.ukhsa-dashboard.data.gov.uk/themes/infectious_disease/sub_themes/gastrointestinal/topics/C-difficile/geography_types/Nation/geographies/England/metrics/c-difficile_cases_countsByOnsetType"

# Add page size
url <- paste0(url, "?page_size=365")

# Set up the request
api_request <- request(url)

# Perform the request
response <- req_perform(api_request)

# Convert to JSON
json_response <- resp_body_json(response)

# Convert to a data frame
results_df <- bind_rows(json_response$results)

# Filter to a single value of stratum
filt <- filter(results_df, stratum == "Total cases")

# Display
gt(filt)
str(filt)
filt$Date<-as.Date(filt$date)
ggplot(filt, aes(
  x = Date,
  y = metric_value
)) +
  geom_line()
```

Plot data
```{r}
library("httr2")
rhinovirus <- read_csv("data/ukhsa-chart-download-rhinovirus.csv")

ggplot(rhinovirus, aes(
  x = date,
  y = metric_value
)) +
  geom_line(color = "red") +
  xlab("Date") +
  ylab("Case counts") +
  ggtitle("Rhinovirus case counts")
rhinovirus <- read_csv("data/ukhsa-chart-download-rhinovirus.csv")
str(rhinovirus)
ggplot(rhinovirus, aes(
  x = date,
  y = metric_value
)) +
  geom_line(color="blue")+
  xlab("Date")+
  ylab("Percentage PCR positive")
```

